### 2023-07-15 template.tex
Update .gitignore, update scripts, add latex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Master Thesis Template by Fabian Schär %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Packages and Document Settings %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[12pt,a4paper,titlepage,oneside,english]{article}

%%% Main Packages %%%
\usepackage[english]{babel}
%\usepackage[ngerman]{babel} % Use this option for German settings.
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

%%% Additional Packages %%%
\usepackage{cite}
\usepackage{framed}
\usepackage{graphicx}
%\usepackage[german]{fancyref}
\usepackage[german,hidelinks]{hyperref} %hidelinks
\usepackage{multirow}
\usepackage[round]{natbib}
\usepackage{setspace}
\usepackage{geometry}
\usepackage{pst-all} % Not working with Sweave!!!

%%% Math Packages %%%
\usepackage{amsmath}
\usepackage{amstext}
\usepackage{amssymb}
\usepackage{theorem}
\usepackage{epsfig}
\usepackage{longtable}

%%% Layout Specifications %%%
\geometry{a4paper, top=35mm, left=40mm, right=40mm, bottom=45mm,
headsep=10mm, footskip=12mm}

%%% Parskip Settings %%%
\setlength{\parskip}{3mm}
\setlength{\parindent}{0mm}

%%% Document Specifications %%%
\title{Template Master's Thesis}
\author{John Doe}

%%%%%%%%%%%%%%%%%%
%%% Title Page %%%
%%%%%%%%%%%%%%%%%%

\begin{document}
%\begin{titlepage}
\begin{center}
\vspace{1em}
\large{Master Thesis}\\
\huge Developing Address Clustering Heuristics for Account-Based Blockchain Networks:\\ An Analysis based on a Specific Address Set \\
\Large \vspace{1em}
Dario Thürkauf
\end{center}

\vspace{1em}
\normalsize
\begin{flushleft}
Supervised by:\\ 
Prof. Dr. Fabian Schär \\
Credit Suisse Asset Management (Schweiz) Professor for \\ 
Distributed Ledger Technologies and Fintech \\
Center for Innovative Finance, University of Basel
\end{flushleft}

\vspace{1em}
\onehalfspacing
\begin{center}
\section*{Abstract}
\end{center}
Assentiar consuetae ha opinionum mentemque ob ii. Ne conflantur de intelligat et me cohibendam. Imaginandi ob to at agnoscerem et mutationum. In methodum ob ii at quicquid lectorum. Procuravi ha dependent ob evidenter tangantur concipere. Immortalem objectivus deo eae rei attingebam ita advertebam quamprimum. Typis patet prius qua nia mem ens. Suppono sim ita pendere nam agnosci quopiam vestiri spondeo dum. Tes illum mundo vetus signa fit talem res his.  \\
\vfill
\textbf{Keywords:} Keyword 1, Keyword 2, Keyword 3, Keyword 4.\\
\noindent\textbf{JEL:} X00, X00, X00
%\end{titlepage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Contents & Declaration%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\pagenumbering{gobble}

\newpage
\pagenumbering{Roman}
\tableofcontents

\vfill
\begin{center}
\includegraphics[width=4cm]{../figures/logo_cif.png}
\end{center}
\newpage
\singlespacing
%\vspace{-1.5cm}
\section*{Declaration of Independent Authorship}
I attest with my signature that I have written this work independently and without outside help. I also attest that the information concerning the sources used in this work is true and complete in every respect. All sources that have been quoted or paraphrased have been marked accordingly. 
Additionally, I affirm that any text passages written with the help of AI-supported technology are marked as such, including a reference to the AI-supported program used. This paper may be checked for plagiarism and use of AI-supported technology using the appropriate software. I understand that unethical conduct may lead to a grade of 1 or ``fail`` or expulsion from the study program.\\

Dario Thürkauf

\begin{figure}[h!]
	\centering
	\hspace{-10cm}
	\includegraphics[width=3cm]{../figures/signature.jpeg}
\end{figure}
%%%%%%%%%%%%%%%%%%%%

\newpage
\onehalfspacing
\pagenumbering{arabic}

%%%%%%%%%%%%%%%%%%%
%%% Introduction%%%
%%%%%%%%%%%%%%%%%%%

\section{Introduction}
Motivation \& Relevance of the topic \\
Literarure Review: Still nascent field, First start with bitcoin/UTXO based, go over to account-based: Victor, Beres et al. \\
Structure of the paper  \\

%%%%%%%%%%%%%%%%%%%%%
%%% Preliminaries %%%
%%%%%%%%%%%%%%%%%%%%%

\section{Preliminaries}

Similar to seminar thesis\\
Account-based blockchains\\
Ethereum, Polygon\\
Smart Contracts, EOAs\\
Transactions, Transfer Events \\
Privacy-enhancing protocols, Tornado \\
Metaverse, Decentraland \\
Token Standards, shorter than in Seminar Thesis, comparable to Metaverse Retailing Paper \\

%%%%%%%%%%%%
%%% Data %%%
%%%%%%%%%%%%

\section{Data Preparation and Retrieval}

Transfer Events, adding Information (isInSet) \\
Transactions \\
Filtering, Intra-set transfers\\
Data Structure, Fields \\
Additional data \\

%%%%%%%%%%%%%%%%%%%%%
%%% Data Analysis %%%
%%%%%%%%%%%%%%%%%%%%%

\section{Data Analysis}
Existing clustering heuristics \\
Graph-based network analysis\\

%%%%%%%%%%%%%%%%%%
%%% Discussion %%%
%%%%%%%%%%%%%%%%%%

\section{Discussion}

%%%%%%%%%%%%%%%%%%
%%% Conclusion %%%
%%%%%%%%%%%%%%%%%%

\section{Conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Literaturverzeichnis %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\setcounter{page}{1}
\pagenumbering{roman}
\onehalfspacing
\addcontentsline{toc}{section}{References}
\bibliography{mybib}
\bibliographystyle{agsm}

%\section{Appendix}
\end{document}

%%% Table
%\begin{table}[h!]
%  \center
%  \begin{tabular}{lcc}
%    \hline\hline
%    Header & Header & Header \\ \hline
%    Entry 1 & $0 \leq x<1$ & $\alpha$\\
%    Entry 2 & $x=1$ & $\beta$\\
%    Entry 3 & $x>1$ & $\gamma$\\
%    \hline\hline
%  \end{tabular}
%  \caption{This is a table}
%  \label{tbl:test}
%\end{table}
%%%